The Implications of Local Fluctuations in the Galactic Midplane for Dynamical Analysis in the \textit{Gaia} Era

### 2019-07-26 paper/ms.tex
fix formatting problem
rest of my comments

\textbf{The true Galactic potential is not strictly axisymmetric, beyond even
the typically-quoted gas clouds, spiral arms, and bar(s). The presence of disk
oscillations has been known for some time, observed in 1957 for H~\textsc{i}
in the outer disk \citep{1957AJ.....62...93K}. The presence of a North-South
asymmetry in stellar number density and velocity in the solar neighborhood has
been found in SDSS data \citep{2012ApJ...750L..41W}, and in velocity in the
LAMOST survey \citep{2013ApJ...777L...5C}, and RAVE data
\citep{2013MNRAS.436..101W}. Beyond the solar neighborhood, low-latitude
overdensities have been observed, e.g., the Monoceros Ring
\citep{2002ApJ...569..245N, 2003MNRAS.340L..21I, 2003ApJ...594L.119C,
2014ApJ...791....9S, 2016ApJ...825..140M}, an overdensity in the direction of
the Triangulum and Andromeda galaxies \citep{2007ApJ...668L.123M,
2014ApJ...793...62S, 2015ApJ...801..105X, 2015MNRAS.452..676P} and A13
\citep{2010ApJ...722..750S, 2017ApJ...844...74L}, and in the direction of the
Galactic Center \citep{2014Natur.509..342F}. There is evidence that these
stellar populations originated from the disk \citep{2018ApJ...854...47S,
2018Natur.555..334B}. See, e.g., \citet{2018MNRAS.481..286L} for a discussion
of these observations.}

@@

Small-scale variations in the density of gas and stars cause the local
midplane position to vary as a function of radius and azimuth. Stars far from
the Sun have a local midplane that differs from our local midplane
extrapolated onto their position. Converting positions and velocities of more
distant stars from a heliocentric to a Galactocentric coordinate system thus
introduces a systematic bias in the $z$ coordinate. We show that this bias
induces non-Gaussian errors in the actions computed for these stars. The
further from the solar neighborhood the target star is, the more likely the
mismatch will result in large systematic uncertainty, especially in the
vertical action $J_z$. A similar argument applies to any remaining uncertainty
in measurements of the Galactocentric radius of the Sun, and to variations in
the LSR.

@@

\textbf{We would like to emphasize that in reality one only ever measures the
position and velocity of a star at a single instance in time, and not over any
significant portion of its orbit. However, if more than one star on the same
orbit can be identified, then it might be possible to constrain the disk
potential in analogy to how stellar streams are used to constrain the Galactic
potential \citep[e.g.][]{1999ApJ...512L.109J, 2001ApJ...551..294I,
2004ApJ...610L..97H, 2015ApJ...801...98S, 2015ApJ...803...80K,
2019MNRAS.483.1427L}. However, identifying the remnants of a disrupted open
cluster (the inner disk equivalent of a halo stellar stream) is a difficult
challenge. One would want to use the same constants of motion that we discuss
are suscept to significant errors. On the other hand, one could use chemical
tagging to identify a disrupted open cluster \citep{2002ARA&A..40..487F}, but
it may be difficult due to the large number of stars that are chemically
indistinguishable but are presumably born in different clusters
\citep{2018ApJ...853..198N}. An alternative technique that would deliver some
of the benefits of finding a disk stream may lie in measuring the
accelerations of stars \citep{2018arXiv181207581S}.}

This example uses an offset in $z$, but analagous effects occur from offsets
in other coordinates, such as the distance to the Galactic center or the LSR.
A similar effect in which actions gain time-dependence due to a time-varying
potential was pointed out by \citet{2015A&A...584A.120B}.

@@

\textbf{Because the energies in the radial and vertical coordinates are related to the
radial and vertical actions by a constant
(Equation~\eqref{eq:JR_Jz_epi_energy}), the fractional error on these two
energies in the as in the actions. The energy in the azimuthal coordinate can
be readily computed, and we therefore conclude that for the given offsets the
error on the total computed energy is
\beq\label{eq:error_total_energy}
\frac{\Delta E}{E} = \frac{\Delta J_R}{J_R} + \frac{\Delta J_{\phi}}{J_{\phi}}
                     + \frac{\Delta J_z}{J_z}\text{.}
\eeq}

@@

We compute the same midplane variation in Figure~\ref{fig:midplane_chervin},
but for four succcessive timesteps of the live N-body simulation of a
Sagittarius encounter \citep{2018MNRAS.481..286L}. Again we have subtracted a
best fit curve of the form given in Equation~(\ref{eq:fit-curve}), with the
values at times $(2.0, 4.0, 6.0, 6.9)\,\Gyr$ being $A = (9.5, 2.5, -210,
-390)\,\pc$, $B = (0.074, 0.039, -36, -57)\,\text{deg}$, and $C = (2.6,-7.8,
-65, -53)\,\pc$. The middle $90\%$ of midplane values across the solar circle
spans $(49, 62, 140, 120)\,\pc$.

>>>>

\textbf{The true Galactic potential is not strictly axisymmetric, beyond even
the typically-quoted gas clouds, spiral arms, and bar(s). The presence of disk
oscillations has been known
%mm for some time, observed in 1957 for
% H~% \textsc{i} in the outer disk \citep{1957AJ.....62...93K}.
    since its observation in H~{sc i} by \citet{1957AJ.....62...93K}.
The presence of a North-South
asymmetry in stellar number density and velocity in the solar neighborhood has
been found in
% mm [should these multiple acronyms be expanded?]
SDSS data \citep{2012ApJ...750L..41W}, and in velocity in the
LAMOST survey \citep{2013ApJ...777L...5C}, and RAVE data
\citep{2013MNRAS.436..101W}. Beyond the solar neighborhood, low-latitude
overdensities have been observed, e.g., the Monoceros Ring
\citep{2002ApJ...569..245N, 2003MNRAS.340L..21I, 2003ApJ...594L.119C,
2014ApJ...791....9S, 2016ApJ...825..140M}, an overdensity in the direction of
the Triangulum and Andromeda galaxies \citep{2007ApJ...668L.123M,
2014ApJ...793...62S, 2015ApJ...801..105X, 2015MNRAS.452..676P} and A13
\citep{2010ApJ...722..750S, 2017ApJ...844...74L}, and in the direction of the
Galactic Center \citep{2014Natur.509..342F}. There is evidence that these
stellar populations originated from the disk \citep{2018ApJ...854...47S,
2018Natur.555..334B}. See, e.g., \citet{2018MNRAS.481..286L} for a discussion
of these observations.}

@@

% mm Small-scale
     Such
variations in the density of gas and stars cause the local
midplane position to vary as a function of radius and azimuth. Stars far from
the Sun have a local midplane that differs from our local midplane
extrapolated onto their position. Converting positions and velocities of more
distant stars from a heliocentric to a Galactocentric coordinate system thus
introduces a systematic bias in the $z$ coordinate. We show that this bias
induces non-Gaussian errors in the actions computed for these stars. The
further from the solar neighborhood the target star is, the more likely the
mismatch will result in large systematic uncertainty, especially in the
vertical action $J_z$. A similar argument applies to any remaining uncertainty
in measurements of the Galactocentric radius of the Sun, and to variations in
the LSR.

@@

%mm [moved para up as it directly follows previous para]
This example uses an offset in $z$, but analagous effects occur from offsets
in other coordinates, such as the distance to the Galactic center or the LSR.
A similar effect in which actions gain time-dependence due to a time-varying
potential was pointed out by \citet{2015A&A...584A.120B}.

\textbf{We would like to emphasize that in reality one only ever measures the
position and velocity of a star at a single instance in time, and not over any
significant portion of its orbit. However, if more than one star on the same
orbit
%mm can be identified, then it might be possible 
     could be identified, then it would be possible
to constrain the disk potential in analogy to how stellar streams are
used to constrain the Galactic  potential
\citep[e.g.][]{1999ApJ...512L.109J, 2001ApJ...551..294I, 
2004ApJ...610L..97H, 2015ApJ...801...98S, 2015ApJ...803...80K,
2019MNRAS.483.1427L}. However, identifying the remnants of a disrupted open
cluster (the inner disk equivalent of a halo stellar stream) is a difficult
challenge. One
%mm would want to use the same constants of motion that we discuss are
%suscept to significant errors. On the other hand, one could use
    possibility would be to use the actions, but as we discuss in this
    work they are subject to substantial errors.  Another possibility would
    be to use
chemical tagging to identify a disrupted open cluster \citep{2002ARA&A..40..487F}, but
%mm it
    this
may be difficult due to the large number of stars that are chemically
indistinguishable
%mm but are
    though
presumably born in different clusters
\citep{2018ApJ...853..198N}. An alternative technique that would deliver some
of the benefits of finding a disk stream may lie in measuring the
accelerations of stars \citep{2018arXiv181207581S}.}

%mm This example uses an offset in $z$, but analagous effects occur from offsets
%in other coordinates, such as the distance to the Galactic center or the LSR.
%A similar effect in which actions gain time-dependence due to a time-varying
%potential was pointed out by \citet{2015A&A...584A.120B}.

@@

\textbf{Because the energies in the radial and vertical coordinates are related to the
radial and vertical actions by a constant
%mm (Equation~\eqref{eq:JR_Jz_epi_energy}),
    (Eq.~[\ref{eq:JR_Jz_epi_energy}])
the fractional error
%mm on
     in
these two
energies
%mm in the as in the actions. The
%[I am guessing as to what this means, so **** check ****]
      evolves in the same way as the actions.  We can see this because the
 energy in the azimuthal coordinate can
be readily computed,
%mm [maybe give it here to ensure that everyone is on the same page?]
%mm and we therefore conclude that
    so
for the given offsets the error
%mm on
    in
the total computed energy is
\beq\label{eq:error_total_energy}
\frac{\Delta E}{E} = \frac{\Delta J_R}{J_R} + \frac{\Delta J_{\phi}}{J_{\phi}}
                     + \frac{\Delta J_z}{J_z}\text{.}
\eeq}

@@

We compute the same midplane variation in Figure~\ref{fig:midplane_chervin},
but for four succcessive timesteps of the live N-body simulation of a
Sagittarius encounter \citep{2018MNRAS.481..286L}. Again we have subtracted a
best fit curve of the form given in Equation~(\ref{eq:fit-curve}), with the
values at times
% mm  [line breaks are needed]
%$(2.0, 4.0, 6.0, 6.9)\,\Gyr$ being $A = (9.5, 2.5, -210,
%-390)\,\pc$, $B = (0.074, 0.039, -36, -57)\,\text{deg}$, and $C = (2.6,-7.8,
%-65, -53)\,\pc$. The middle $90\%$ of midplane values across the solar circle
%spans mm $(49, 62, 140, 120)\,\pc$.
    (2.0, 4.0, 6.0, 6.9)~Gyr being $A$ = (9.5, 2.5, -210, -390)~pc,
    $B$ = (0.074, 0.039, -36, -57)~deg, and $C$ = (2.6,-7.8, -65,
    -53)~pc. The middle $90\%$ of midplane values across the solar
    circle spans (49, 62, 140, 120)~pc.
